---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# harp <a href=#><img src='man/figures/harp_logo_dark.svg' align="right" height="131.5" style="margin-left:30px" /></a>


{_harp_} is a meta-package that attaches functionality from the [{_harpIO_}](https://harphub.github.io/harpIO), [{_harpPoint_}](https://harphub.github.io/harpPoint), [_{harpVis_}](https://harphub.github.io/harpVis) and [{_harpSpatial_}](https://harphub.github.io/harpSpatial) packages to your session.  

## Installation
```{r eval=FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("harphub/harp")
```

### Installing on ecgate
Before installing on ecgate you need to load the proj4 and R modules before starting R and executing the above:
```{bash eval=FALSE}
module load proj4
module load R
```

ecgate has had an issue with the {_RSQLite_} package that provides functionality for reading from and writing to SQLite files. If you experience problems with RSQLite during installation, this can be alleviated by installing an older version of {_RSQLite_} before attempting to install {_harp_} using `remotes::install_version`:
```{r eval=FALSE}
remotes::install_version("RSQLite", "2.2.5")
```

You may be asked if you want to update versions of certain packages. In general it is best to answer with the number that corresponds to All. However, when {_harpIO_} begins to install, you need to input the numbers (separated by spaces) for all packages excluding RSQLite. 

### System libraries
The [{_harpIO_}](https://harphub.github.io/harpIO) and [{_harpSpatial_}](https://harphub.github.io/harpSpatial) packages use the [{_meteogrid_}](https://github.com/harphub/meteogrid) package for interpolation and the geogrid class for storing gridded data. This package makes use of the [PROJ4](https://proj4.org) library for handling projections. If you do not have PROJ4 installed you can install with:

```{bash eval=FALSE}
sudo apt-get install libproj-dev
```

If you do not have sudo rights, try installing a user local version, or speak to your system administrator. 

If you have the PROJ4 libraries installed in a standard location (e.g. /usr/local) {_meteogrid_} will install without problems. However, if the PROJ4 libraries are in a non standard location, you need to tell the install function where they are:
```{r eval=FALSE}
remotes::install_github(
  "harphub/harpIO",
  configure.args = c(
    meteogrid = "--with-proj=/path/to/proj"
  )
)
```

Alternatively you can set environment variables
```{bash eval=FALSE}
export PROJ4_DIR=/path/to/proj
```

If you include these environment variables in your .bashrc file, or equivalent, you won't need to worry about it when you wish to install an update to meteogrid.

Or you can set compiler and linker options in the file $HOME/.R/Makevars
```{bash eval=FALSE}
CPPFLAGS=-I/path/to/proj/include
LDFLAGS=-L/path/to/proj/lib -Wl,-rpath,/path/to/proj/lib
```

In this case you only have to set them once and not worry about it when you wish to install an update to {_meteogrid_}.

When setting environment variables or creating a Makevars file, R must be restarted for the changes to take effect before running `remotes::install_github("harphub/harp")`.

## Workflows
The purpose of this website is to demonstrate common workflows, such as point and spatial verification, score card generation, and data analysis using the {_harp_} packages. 

## harp book
An online [book](https://harphub.github.io/harp_tutorial) that collects together tutorials on how to use harp is currently being worked on. Feel free to make suggestions, or to contribute by filing issues or pull requests [here](https://github.com/harphub/harp_tutorial) 
